# Using synapsis

# Setup in Rstudio

### installing library from gitlab

To use synapsis, you will need the following packages:

- stats
- EBImage
- graphics
- utils

And to run this tutorial, you will need:

- tidyverse
- ggplot2
- knitr
- rmarkdown

```{r}
#if (!requireNamespace("BiocManager", quietly = TRUE))
#install.packages("BiocManager")
#BiocManager::install("EBImage")
# change to synapsis eventually

```

```{r}
#devtools::install_git('https://gitlab.svi.edu.au/lmcneill/synapsis')
```

### loading synapsis

```{r}
library(synapsis)
```
### checking documentation

```{r}
??count_foci
```

## Data preparation

Please download test1.zip and test2.zip in the "coding" chat.
Double click to unzip. Drag the folders into the place you'll be working from, and make note of the path.
For example, if I want to put them in a folder I created called "test-data-all" in "imaging", in "svi", in the "Documents" folder, my path looks like:

```{r}
path = "~/Documents/svi/imaging/test-data-all"
```

## might need these libraries?

```{r}
#library(knitr)
#library(rmarkdown)
#library(EBImage)
#library(ggplot2)
```

# Calling functions on data

## Cropping routine. 

*cell_area variables are based on our mouse imaging data. For different species with larger/ smaller nuclei or other teams using different magnifications, we would need to change this (with some finetuning, which is made easier by rknit)

There is an annotation setting, which we would switch to "on"
```{r}
#auto_crop(path, annotation = "on", max_cell_area = 30000, min_cell_area = 7000)
```
Here we called path, plus other optional parameters (that would otherwise take on default values). But only path is essential. This is because auto_crop has built-in default values which are assumed when the user doesn't specify.

*Comment this auto_crop chunk out now* if a crops folder with three channels per image was successfully generated. Now that cell candidates have been cropped, we don't need/want to wait around for that again!

## Getting pachytene 

```{r}
SYCP3_stats <- get_pachytene(path,ecc_thresh = 0.8, area_thresh = 0.04)
```
SYCP3_stats is a data frame summarising some features of the cells classified as pachytene.


## Counting foci

```{r}
foci_counts <- count_foci(path,offset_factor = 3, brush_size = 3, brush_sigma = 3, annotate = "on")
```
Make sure every line prior to the above chunk is commented out, because we want to knit with annotate = "on" to check that synapsis is counting close or same as a manual count.

If it's identifying too many things as focis, try increasing some of the input parameters like offset_factor or brush_size.
If it's not identifying any foci but there are clearly foci there, try decreasing those parameters.
# Statistics

## some basic statistics
```{r}
### comparing groups
counts <- foci_counts$foci_count
hist(as.numeric(counts))
counts_mod <- foci_counts[as.numeric(foci_counts$foci_count) > 0,]
counts_mod <- foci_counts[as.numeric(foci_counts$foci_count) < 40,]
#counts_mod <- counts_mod[as.numeric(counts_mod$percent_on) > 0.55,]
# counts_mod <- counts_mod[as.numeric(counts_mod$sd_foci) <20,]
counts <- counts_mod$foci_count
hist(as.numeric(counts))
counts_KO <- counts_mod[counts_mod$genotype == "Fancm-/-",]
counts_WT <- counts_mod[counts_mod$genotype == "Fancm+/+",]

count_KO <- counts_KO$foci_count
count_WT <- counts_WT$foci_count
mean(as.numeric(count_KO), na.rm= TRUE)
mean(as.numeric(count_WT), na.rm= TRUE)
sd(as.numeric(count_KO), na.rm= TRUE)
sd(as.numeric(count_WT), na.rm= TRUE)

c1 <- rgb(173,216,230,max = 255, alpha = 140, names = "lt.blue")
c4 <- rgb(255,200,50, max = 255, alpha = 120, names = "lt.orange")
A <- hist(as.numeric(count_WT),plot = FALSE)
B <-  hist(as.numeric(count_KO), plot = FALSE )
plot(A,ylim = c(0,40),  main = "Pachytene", col = c4, xlab = "foci count") 
plot(B, col = c1, add = TRUE) 
```


## comparison testing

### anova test
```{r}
## anova test
counts_mod$group <- factor(counts_mod$genotype, c("Fancm-/-", "Fancm+/+"))
outfit <- lm(foci_count ~ genotype, data=counts_mod)
outfit
#df.residual(outfit)
#sigma(outfit)
#model.matrix(outfit)
outfit0 <- lm(foci_count ~ 1, data=counts_mod)
anova(outfit0, outfit)
```
 
### t test
```{r}
t.test(as.numeric(count_KO),as.numeric(count_WT))
```

#boxplot
```{r}
library(tidyverse)
counts_mod %>% 
  ggplot(aes(x=genotype, y=as.numeric(foci_count), fill=genotype)) + 
  geom_boxplot(width=0.5,lwd=1.5) +
  geom_jitter(width=0.15)+
  labs(subtitle="MLH3 foci counts")
```

The next function takes a long time so has been commented out.
```{r}
#source("measure_distances.R")
#df_dist <- measure_distances(path, annotate = "off")

```
