# Using synapsis

---
title: "Using synapsis"
author: "Lucy McNeill"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    
    
    number_sections: true
    toc: yes
    toc_float: yes
---

# Setup in Rstudio

### installing library from gitlab

To use synapsis, you will need the following packages:

- stats
- EBImage
- graphics
- utils

And to run this tutorial, you will need:

- tidyverse
- ggplot2
- knitr
- rmarkdown


```{r, eval = FALSE}
devtools::install_git('https://gitlab.svi.edu.au/lmcneill/synapsis')
```

### loading synapsis

```{r}
library(synapsis)
```
### checking documentation

## Data preparation

Please download test1.zip and test2.zip in the "coding" chat.
Double click to unzip. Drag the folders into the place you'll be working from, and make note of the path.
For example, if I want to put them in a folder I created called "test-data-all" in "imaging", in "svi", in the "Documents" folder, my path looks like:

```{r}
path = "~/Documents/svi/imaging/test-data-all"
```
# Calling functions on data

## Cropping routine. 

You can type 
```{r,  eval = FALSE}
??auto_crop_fast
```

There is an annotation setting, which we switch to "on".  max_cell_area and  min_cell_area have been calibrated to our data set, where the subject are mouse cells and magnification kept constant. Let's run it on the first five images by setting test_amount = 5
```{r}
auto_crop_fast(path, annotation = "off", max_cell_area = 30000, min_cell_area = 7000, test_amount = 4)
```
Here we called path, plus other optional parameters (that would otherwise take on default values). But only path is essential. This is because auto_crop_fast has built-in default values which are assumed when the user doesn't specify.

A crops folder with three channels per "viable cell" should have been generated inside the folder where these images are kept i.e. in path.

## Getting pachytene 

```{r}
SYCP3_stats <- get_pachytene(path,ecc_thresh = 0.8, area_thresh = 0.04, annotation = "on")
```
SYCP3_stats is a data frame summarising some features of the cells classified as pachytene.

## Bigger data set.

Now we will run auto_crop on the bigger data set (silently/ without annotation) to do some significance testing.
```{r, echo = FALSE}
auto_crop_fast(path, annotation = "off", max_cell_area = 30000, min_cell_area = 700)
```


## Counting foci

Now let's count the foci for each genotype.

```{r, eval}
SYCP3_stats <- get_pachytene(path,ecc_thresh = 0.8, area_thresh = 0.04, annotation = "off")
foci_counts <- count_foci(path,offset_factor = 3, brush_size = 3, brush_sigma = 3, annotation = "off",stage = "pachytene")
```
Make sure every line prior to the above chunk is commented out, because we want to knit with annotate = "on" to check that synapsis is counting close or same as a manual count.

If it's identifying too many things as focis, try increasing some of the input parameters like offset_factor or brush_size.
If it's not identifying any foci but there are clearly foci there, try decreasing those parameters.
# Statistics

## some basic statistics
```{r}
### comparing groups
counts <- foci_counts$foci_count
counts_mod <- foci_counts[as.numeric(foci_counts$foci_count) > 0,]
counts_mod <- foci_counts[as.numeric(foci_counts$foci_count) < 40,]
#counts_mod <- counts_mod[as.numeric(counts_mod$percent_on) > 0.55,]
# counts_mod <- counts_mod[as.numeric(counts_mod$sd_foci) <20,]
counts <- counts_mod$foci_count
counts_KO <- counts_mod[counts_mod$genotype == "Fancm-/-",]
counts_WT <- counts_mod[counts_mod$genotype == "Fancm+/+",]

count_KO <- counts_KO$foci_count
count_WT <- counts_WT$foci_count
mean(as.numeric(count_KO), na.rm= TRUE)
mean(as.numeric(count_WT), na.rm= TRUE)
sd(as.numeric(count_KO), na.rm= TRUE)
sd(as.numeric(count_WT), na.rm= TRUE)

c1 <- rgb(173,216,230,max = 255, alpha = 140, names = "lt.blue")
c4 <- rgb(255,180,50, max = 255, alpha = 120, names = "lt.orange")
A <- hist(as.numeric(count_WT),plot = FALSE)
B <-  hist(as.numeric(count_KO), plot = FALSE )
plot(A,ylim = c(0,20),  main = "Pachytene", col = c4, xlab = "foci count per cell") 
plot(B, col = c1, add = TRUE) 
```


## comparison testing

### anova test
```{r}
## anova test
counts_mod$group <- factor(counts_mod$genotype, c("Fancm-/-", "Fancm+/+"))
outfit <- lm(foci_count ~ genotype, data=counts_mod)
outfit
#df.residual(outfit)
#sigma(outfit)
#model.matrix(outfit)
outfit0 <- lm(foci_count ~ 1, data=counts_mod)
anova(outfit0, outfit)
```
 
### t test
```{r}
t.test(as.numeric(count_KO),as.numeric(count_WT))
```
now that we have a p value we can paste this on the histogram 
```{r}
c1 <- rgb(173,216,230,max = 255, alpha = 140, names = "lt.blue")
c4 <- rgb(255,180,50, max = 255, alpha = 120, names = "lt.orange")
A <- hist(as.numeric(count_WT),plot = FALSE)
B <-  hist(as.numeric(count_KO), plot = FALSE )
plot(A,ylim = c(0,20),  main = "Pachytene", col = c4, xlab = "foci count per cell") 
text(x = 10, y = 15, label = "anova p value = 0.01*", col = "black", cex = 1)
plot(B, col = c1, add = TRUE) 
```

#boxplot
```{r}
library(tidyverse)
counts_mod %>% 
  ggplot(aes(x=genotype, y=as.numeric(foci_count), fill=genotype)) + 
  geom_boxplot(width=0.5,lwd=1.5) +
  geom_jitter(width=0.15)+
  labs(subtitle="MLH3 foci counts")
```

We won't run this for the notebook yet: 
```{r, eval = FALSE}
df_dist <- measure_distances(path, annotation = "off")
print(df_dist)

```
```{r, eval = FALSE}
pass_only <- df_dist[df_dist$pass_fail == "pass",]

distances <- pass_only$fractional_distance
distances_KO <- pass_only[pass_only$genotype == "Fancm-/-",]
distances_WT <- pass_only[pass_only$genotype == "Fancm+/+",]

distance_KO <- distances_KO$fractional_distance
distance_WT <- distances_WT$fractional_distance
mean(as.numeric(distance_KO), na.rm= TRUE)
mean(as.numeric(distance_WT), na.rm= TRUE)
sd(as.numeric(distance_KO), na.rm= TRUE)
sd(as.numeric(distance_WT), na.rm= TRUE)

c1 <- rgb(173,216,230,max = 255, alpha = 140, names = "lt.blue")
c4 <- rgb(255,180,50, max = 255, alpha = 120, names = "lt.orange")
A <- hist(as.numeric(distance_WT),plot = FALSE)
B <-  hist(as.numeric(distance_KO), plot = FALSE )
plot(A,ylim = c(0,9),xlim = c(0,1),  main = "Pachytene", col = c4, xlab = "foci distance as fraction of total length") 
text(x = 0.2, y = 7, label = "anova p value 0.9 (NS)", col = "black", cex = 1)
plot(B, col = c1, add = TRUE) 
```
```{r, eval = FALSE}
t.test(as.numeric(distance_KO),as.numeric(distance_WT))
```
```{r, eval = FALSE}
column <- as.data.frame(count_WT)
write.csv(column, "column4.csv",row.names = FALSE)

```
